---
title: "Pixels & Points"
author: "Ben Ortiz Ulloa"
date: "1/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

## Raster Basics

### Create Matrix

```{r}
intro_mat <- matrix(1:25, ncol = 5)
intro_mat
```

### Create RAster

```{r}
intro_rast <- raster::raster(intro_mat)
intro_rast
```

### Plot raster

```{r}
raster::plot(intro_rast)
```

### Matrix and rasters can store booleans

```{r}
intro_mat_bool <- rbind(
  c(F,F,T,F,F),
  c(F,F,T,F,F),
  c(T,T,T,T,T),
  c(F,F,T,F,F),
  c(F,F,T,F,F)
)

intro_mat_bool
```

### as a raster booleans are stored as 0 and 1

```{r}
intro_rast_bool <- raster::raster(intro_mat_bool)
intro_rast_bool
```

```{r}
raster::plot(intro_rast_bool)
```

### Rasters booleans can be used to perform operations on truth pixels

```{r}
intro_bool_remove = intro_rast
intro_bool_remove[intro_rast_bool] <- NA
raster::as.matrix(intro_bool_remove)
```

### NA values are plotted as transprent

```{r}
raster::plot(intro_bool_remove)
```

### rasters can use themselves to perform queried operations

```{r}
intro_rast_query <- intro_rast
intro_rast_query[intro_rast_query %% 2] <- NA

raster::plot(intro_rast_query)
```

## Lets do this with real data

```{r}
v <- raster::raster(volcano)

raster::plot(v)
```

### col argument makes it easier to change colors

```{r}
gray.colors(10)
```

```{r}
raster::plot(v, col = gray.colors(10))
```

### an alternative to just coloring is to create contours

```{r}
x <- raster::rasterToContour(v, nlevels = 20) %>%
  sf::st_as_sf()
  

plot(x)
```

### contours aren't very helpful...

```{r}
x %>%
  dplyr::mutate(level = level %>% as.character() %>% as.numeric()) %>%
  dplyr::filter(level >= 130) %>%
  dplyr::mutate(geometry = sf::st_cast(geometry, 'MULTIPOLYGON'),
                inundated = level <= 150) %>%
  dplyr::select(inundated) %>%
  plot(pal = c('green', 'blue'))
```

## RAYSHADER!!!!

### rayshader is often used to create better coloring options for elevation maps

```{r}
library(rayshader)

# t - transpose - is necessary to convert the matrix to a layout that rayshader likes
volcanoshadow = sphere_shade(t(volcano))

rayshader::plot_map(volcanoshadow)
```

### Simply put the plot into sphere shaded image into plot (mesh)

```{r}
rayshader::plot_3d(hillshade = volcanoshadow, heightmap = t(volcano))

rayshader::render_snapshot()
```


## But what else can we do?


### images are genrally stored as multicolor arrays

```{r}
# rgb values can be stored as an umber from 0 to 255

set.seed(4321)
intro_arr <- round(array(runif(75, 1, 255), c(5,5,3)))
intro_arr
```

### rasters only can contain one value per pixel, bricks can hold multiple

```{r}
intro_brick <- raster::brick(intro_arr)
names(intro_brick) <- c('red', 'green', 'blue')
intro_brick
```

### use plotrgb to visualize 3 different values from a brick

```{r}
raster::plotRGB(intro_brick, r = 1, g = 2, b = 3)
```

### bricks can also be queried on each layer

```{r}
intro_brick_bool <- intro_brick[[1]] %% 2 & intro_brick[[2]] %% 2 & intro_brick[[3]] %% 2
raster::plot(intro_brick_bool)
```

```{r}
intro_brick_query <- intro_brick
intro_brick_query[intro_brick_bool] <- NA
raster::plotRGB(intro_brick_query)
```



```{r}
dvdc <- raster::brick('../Assets/data-viz-dc.png')

#fourth layer is typically alpha or transparency
dvdc
```

```{r}
raster::plotRGB(dvdc, 1, 2, 3)
```

### aggregate nearby pixels to get summary statistics - also lowers resolution

```{r}
dvdc_smallAgg <- raster::aggregate(dvdc, fact = 9, fun = mean)
raster::plotRGB(dvdc_smallAgg)
```

### it's important to know how big of an area you want to care about

```{r}
dvdc_largeAgg <- raster::aggregate(dvdc, 45, mean)
raster::plotRGB(dvdc_largeAgg)
```

### There are only two colors, black and white

```{r}
### should be 0, but some grays in images so we'll do less than 50
dvdc_black <- dvdc[[1]] <= 50 & dvdc[[2]] <= 50 & dvdc[[3]] <= 50
dvdc_black
raster::plot(dvdc_black)
```

```{r}
#should be 255, but some grays so >= 240
dvdc_white <- dvdc[[1]] >= 240 & dvdc[[2]] >= 240 & dvdc[[3]] >= 240
dvdc_white
raster::plot(dvdc_white)
```

```{r}
dvdc_cover <- dvdc
dvdc_cover[dvdc_black] <- NA
raster::plotRGB(dvdc_cover, colNA = 'purple')
```

## let's fill the na values with an other image

```{r}
amurika <- raster::brick('../Assets/american-flag-2260839_960_720.jpg')
raster::plotRGB(amurika)
```

## Raster Operations Require Identical Dimensions

```{r}
raster::extent(dvdc_cover)
raster::extent(amurika)
```

```{r}
dim(dvdc_cover)
dim(amurika)
```

```{r}
raster::res(dvdc_cover)
raster::res(amurika)
```

### Method I: Crop & Extend

```{r}
# add an alpha layer to flag
amurika[[4]] <- 255
small_amurika <- raster::crop(amurika, dvdc_cover)
small_amurika <- raster::extend(small_amurika, dvdc_cover)
raster::plotRGB(small_amurika, colNA = 'purple')
```

```{r}
dvdc_amurika <- raster::cover(dvdc_cover, small_amurika)
raster::plotRGB(dvdc_amurika)
```

### Method II:  resample and cover

```{r}
long_amurika <- amurika
raster::extent(long_amurika) <- raster::extent(dvdc_cover)
raster::plotRGB(long_amurika)
```

```{r}
raster::res(long_amurika)
```

```{r}
long_amurika_res <- raster::resample(long_amurika, dvdc_cover, 'ngb')
raster::res(long_amurika_res)
```

```{r}
raster::plotRGB(long_amurika_res)
```

```{r}
dvdc_amurika2 <- raster::cover(dvdc_cover, long_amurika_res)
raster::plotRGB(dvdc_amurika2)
```

## Geospatial Rasters

```{r}
dc <- raster::brick('../Assets/m_3807708_se_18_1_20160718_20160928.jp2')

dc
```

```{r}
raster::plotRGB(dc)
```



```{r}
# c(38.8998515, -77.0318053)
wh_extent <- rbind(c(-77.03, 38.894383),
                       c(-77.03969, 38.90232)) %>%
  sp::SpatialPoints(proj4string = sp::CRS("+init=epsg:4326")) %>%
  sp::spTransform(sf::st_crs(dc)$proj4string) %>%
  raster::extent()

wh_extent
```


```{r}
wh <- raster::crop(dc, wh_extent)
raster::plotRGB(wh)
```

```{r, results=F}
png('../Assets/almost_white_house.png', dim(wh)[2], dim(wh)[1])
raster::plotRGB(wh)
dev.off()
```

```{r}
marked_wh <- raster::brick('../Assets/almost_white_house_marked.png')
marked_bool <- marked_wh[[1]] == 163 & marked_wh[[2]] == 73 & marked_wh[[3]] == 164
marked_bool[marked_bool == 0] <- NA
raster::extent(marked_bool) <- raster::extent(wh)
masked_wh <- raster::mask(wh, marked_bool)
raster::plotRGB(masked_wh)
```

```{r}
wh_bbox <- !is.na(masked_wh)
wh_bbox[!wh_bbox] <- NA
wh_bbox <- wh_bbox[[1]] %>%
  raster::rasterToPolygons() %>% 
  raster::extent()

wh_zoom <- raster::crop(wh, wh_bbox)
raster::plotRGB(wh_zoom)
```

```{r}
crashes <- sf::read_sf('../Assets/Crashes_in_DC/Crashes_in_DC.shp')
```


```{r}
wh_area <- wh %>%
  sf::st_bbox() %>%
  sf::st_as_sfc() %>%
  sf::st_as_sf()

wh_crashes <- crashes %>%
  sf::st_transform(sf::st_crs(wh_area)) %>%
  dplyr::filter(sf::st_intersects(., wh_area, F))
```


```{r}
wh_gray <- wh[[1]] + wh[[2]] + wh[[3]]

raster::plot(wh_gray, col = gray.colors(100))
plot(wh_crashes$geometry, add = T, pch = 21, bg = 'red', cex =.5)
```


```{r}
crash <- raster::rasterize(wh_crashes, wh, field = 1, fun = function(x,...) length(x))

raster::plot(crash, col = rev(heat.colors(50)))
```

```{r}
crash_agg <- raster::aggregate(crash, 20) 

raster::plot(crash_agg, col = rev(heat.colors(50)))
```


```{r}
danger_zone <- crash_agg >= quantile(raster::values(crash_agg), 
                                               .95, 
                                               na.rm = T)

crash_agg[!danger_zone] <- NA

raster::plot(wh_gray, col = gray.colors(100), legend = F)
raster::plot(crash_agg, add = T, col = rev(heat.colors(5)))

```

```{r}
crash_shape <- crash_agg %>%
  raster::rasterToPolygons() %>%
  sf::st_as_sf()

raster::plot(wh_gray, col = gray.colors(100), legend = F)
plot(crash_shape, add = T)
```

```{r}
highest_crash_area <- crash_shape %>%
  dplyr::arrange(desc(layer)) %>%
  head(1)

crashes_in_area <- wh_crashes %>% 
  dplyr::filter(sf::st_intersects(., highest_crash_area, F))

head(crashes_in_area)
```

```{r}
wh_crop <- raster::crop(wh_gray, raster::extent(highest_crash_area))

raster::plot(wh_crop, col = gray.colors(50), legend =F)
plot(crashes_in_area$geometry, add = T, pch = 21, bg = 'red', alpha = .1)
```

